False Discovery Rates, A New Deal

### 2016-04-16 talks/one_hour_seminar.Rmd
update seminar

---
title: "False Discovery Rates, A New Deal"
author: Matthew Stephens
date: March 9, 2015
output: ioslides_presentation
---

@@

- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
- Great for sharing material across multiple people (eg student and advisor!)
- An amateur example: [http://github.com/stephens999/ash](http://github.com/stephens999/ash)

@@

## Commercial Break:

A quick ad for a recently-started project I call "Dynamic Statistical Comparisons":

- Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods.

- How? Centralize reproducible statistical comparisons in easily-extensible public repositories.

## Dynamic Statistical Comparisons: advantages

- Authors can run their own methods. 
- Ease of adding new data can quickly uncover over-tuning to specific datasets.
- Work is divided, and not duplicated.
- Promotes collaboration through competition. 
- Ensures results are reproducible, and extendable.

## More on git, github, knitr, reproducibility, DSCs

- Google "The git book", to get started on git.
- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
- Google "donohoe buckheit" for "Wavelab and reproducible research"
- [http://stephens999.github.io/blog/](http://stephens999.github.io/blog/) for more on DSC

@@

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (EM algorithm).

- By allowing $K$ large, and $\sigma_k$ to span a dense grid of values,
we get a flexible unimodal symmetric distribution.

@@

- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be ``shrunk" towards the overall mean (0).

@@

## Simulated example: $\pi_0$ not identifiable.

```{r,echo=FALSE,include=FALSE}
load("../paper/Rcode/sim1.RData") #load simulation results
source("../paper/Rcode/plot_pi0.R")
```

```{r, echo=FALSE,fig.height=4,fig.cap=""}
plot_pi0(list(simres1a))
```

## Simulated Example: so fdr not identifiable

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
source("../paper/Rcode/plot_lfsr.R")
  plot_lfsr(list(simres1a),0.1,ptype="lfdr")
```

## Simulated Example: fsr much more identifiable

```{r, echo=FALSE,fig.height=4,fig.cap="",warning=FALSE,message=FALSE}
  plot_lfsr(list(simres1a),0.1,ptype="lfsr")
```

@@

## Other Applications

- Widely applicable: requiring only an estimated
effect size and standard error for each object.

- E.g. Currently applying it to wavelet shrinkage applications.

## Next steps?

- Incorporate shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al)

- Allow $g(\cdot;\pi)$ to depend on covariates $X$.

- Allow for correlations in the measured $\hat\beta_j$.

@@

- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Scott Powers, Sen Tian, Wei Wang, Zhengrong Xing.

>>>>

---
title: "False Discovery Rates, A New Deal"
author: Matthew Stephens
date: April 15, 2016
output: ioslides_presentation
---

@@

- A repository: a central place in which an aggregation of data is kept and maintained in an organized way (searcharticle.com)
- Great for sharing material across multiple people (eg student and advisor!)
- For example: [https://github.com/jdblischak/singleCellSeq](https://github.com/jdblischak/singleCellSeq)
- A template (in progress): [https://github.com/jhsiao999/ashlar](https://github.com/jhsiao999/ashlar)

@@

<!-- ## Commercial Break:  -->

<!-- A quick ad for a recently-started project I call "Dynamic Statistical Comparisons": -->

<!-- - Overall Goal: Improve how the statistics and genomics (and other) communities make comparisons among methods. -->

<!-- - How? Centralize reproducible statistical comparisons in easily-extensible public repositories. -->

<!-- ## Dynamic Statistical Comparisons: advantages -->

<!-- - Authors can run their own methods.  -->
<!-- - Ease of adding new data can quickly uncover over-tuning to specific datasets. -->
<!-- - Work is divided, and not duplicated. -->
<!-- - Promotes collaboration through competition.  -->
<!-- - Ensures results are reproducible, and extendable. -->

## More on git, github, knitr, reproducibility

- Google "The git book", to get started on git.
- Google "Karl Broman github tutorial" for statistics-oriented intro to github.
- Google "donohoe buckheit" for "Wavelab and reproducible research"

@@

- Estimating $g$ comes down to estimating $\pi$. Joint estimation of $\pi_0,\pi$ easy by maximum likelihood (convex optimization).

- Setting $K$ large, with dense grid of $\sigma_k$, yields flexible unimodal symmetric distribution.

@@

- Because $g(\beta)$ is unimodal, the point estimates (and CIs) will tend to be "shrunk" towards the overall mean (0).

@@

## More extensive simulations

![Image](../paper/figures/scenarios_density.pdf)

$$\beta_j \sim \pi_0 \delta_0(.) + (1-\pi_0) g(.)$$

Goals: Estimate $\pi_0$? Estimate $g$?

## Estimation of $\pi_0$

![Image](../paper/figures/pi0_est.pdf)

- All methods are, generally, "conservative" (deliberately).

- `ash` generally least conservative.

## Estimation of $g$

- Estimating $g$ is a “deconvolution problem”.
- Notoriously difficult in general. 
- Efron emphasises the difficulties of implementing a stable general algorithm:

<div class="red">
“the effort foundered on practical difficulties involving the perils of deconvolution... Maybe I am trying to be overly nonparametric... but it is hard to imagine a generally satisfactory parametric formulation...”
</div>

## Unimodal assumption stabilizes estimation of $g$

![Image](../paper/figures/egcdf.pdf)

@@

## Applications/Generalizations

- The idea of inducing sparsity/shrinkage via unimodal distributions is very general!

- smash: Smoothing via adaptive shrinkage (wavelet-based)

- vash: shrinkage of variances and not just means. (e.g. ``moderated $t$ test", Smyth et al) [Mengyin Lu]

- flash: shrinkage of factors and loadings in factor analysis [Wei Wang; David Gerard]

- mash: allow for matrix of observations/effects $\hat\beta_{ij}$ [Sarah Urbut, Gao Wang]

- cash: allow $g(\cdot;\pi)$ to depend on covariates $X$ [Scott Powers, Lei Sun]

@@

- Including  Chaoxing Dai, Mengyin Lu, Ester Pantaleo, Sen Tian, Wei Wang, Zhengrong Xing, Sarah Urbut
